---
title: "Black-Scholes Option Pricing in R"
output:
  html_document:
    toc: true
---

## Introduction

The Black-Scholes model is a mathematical model used to determine the price of European call and put options. This document demonstrates how to implement this model using R.

## R Code for Black-Scholes Calculation

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Function Definitions

We start by defining helper functions for calculating the cumulative distribution function of the standard normal distribution ($N$), and the intermediate values $d_1$ and $d_2$.

```{r black-scholes-functions}
# Load necessary library
library(stats)

# Function to calculate N(x)
N <- function(x) {
  pnorm(x)
}

# Function to calculate d1 and d2
d1 <- function(S0, X, r, sigma, T) {
  (log(S0 / X) + (r + sigma^2 / 2) * T) / (sigma * sqrt(T))
}

d2 <- function(S0, X, r, sigma, T) {
  d1(S0, X, r, sigma, T) - sigma * sqrt(T)
}
```

### Black-Scholes Call Option Price

The call option price $C$ is calculated using the formula:

$$ C = S_0 N(d_1) - X e^{-rT} N(d_2) $$

```{r black-scholes-call}
# Function to calculate Black-Scholes call option price
black_scholes_call <- function(S0, X, r, sigma, T) {
  C <- S0 * N(d1(S0, X, r, sigma, T)) - X * exp(-r * T) * N(d2(S0, X, r, sigma, T))
  return(C)
}
```

### Black-Scholes Put Option Price

The put option price $P$ is calculated using the formula:

$$ P = X e^{-rT} N(-d_2) - S_0 N(-d_1) $$

```{r black-scholes-put}
# Function to calculate Black-Scholes put option price
black_scholes_put <- function(S0, X, r, sigma, T) {
  P <- X * exp(-r * T) * N(-d2(S0, X, r, sigma, T)) - S0 * N(-d1(S0, X, r, sigma, T))
  return(P)
}
```

### Example Usage

We will use the Black-Scholes functions to calculate the prices of a call and put option with the following parameters:

- $S_0 = 100$ (Current stock price)
- $X = 100$ (Strike price)
- $r = 0.05$ (Risk-free interest rate)
- $\sigma = 0.2$ (Volatility)
- $T = 1$ (Time to maturity)

```{r example-usage}
# Example usage
S0 <- 100  # Current stock price
X <- 100   # Strike price
r <- 0.05  # Risk-free interest rate
sigma <- 0.2  # Volatility
T <- 1     # Time to maturity (in years)

call_price <- black_scholes_call(S0, X, r, sigma, T)
put_price <- black_scholes_put(S0, X, r, sigma, T)

print(paste("Black-Scholes Call Option Price:", call_price))
print(paste("Black-Scholes Put Option Price:", put_price))
```

### Output

```{r output}
S0 <- 100  # Current stock price
X <- 100   # Strike price
r <- 0.05  # Risk-free interest rate
sigma <- 0.2  # Volatility
T <- 1     # Time to maturity (in years)

call_price <- black_scholes_call(S0, X, r, sigma, T)
put_price <- black_scholes_put(S0, X, r, sigma, T)

cat("Black-Scholes Call Option Price:", call_price, "\n")
cat("Black-Scholes Put Option Price:", put_price, "\n")
```

## Conclusion

The Black-Scholes model is a powerful tool for pricing European options. This document has demonstrated how to implement the model in R and provided an example usage of the code.